Characteristic polynomials of p-adic matrices

### 2017-02-04 charpoly/charpoly.tex
formating for cond

\newcommand{\noopsort}[1]{}
\DeclareMathOperator{\NP}{NP}
\DeclareMathOperator{\HP}{HP}
\DeclareMathOperator{\PP}{PP}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\com}{Com}
\DeclareMathOperator{\Grass}{Grass}
\DeclareMathOperator{\Lat}{Lat}
\DeclareMathOperator{\round}{round}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\rec}{rec}

@@

\begin{prop}
Let $cond(M)$ be the valuation of the last invariant factor of $M.$
If $dM$ is a flat precision of $O(\pi^N)$ on $M$ then $M^{-1}$
can be computed at precision $O(\pi^{N-2cond(M)})$ by a \textbf{SNF} computation
and this lower-bound is optimal
at least when $N$ is large.
\end{prop}
\begin{proof}
The smallest valuation of a coefficient of $M^{-1}$ is $-cond(M).$
It is $-2cond(M)$ for $M^{-2}$ and it is then clear that $-2cond(M)$
can be obtained as the valuation of a coefficient of $du(M) \cdot dM$
and the smallest that can be achieved this way for $dM$ in a precision lattice
of flat precision. Hence the optimality of the bound given, at least when 
$N$ is large, thanks to Lemma 3.14 of \cite{caruso-roe-vaccon:14a}

Now, the computation of an SNF form over a CDVF was described in \cite{Vaccon-these}.
From $M$ known at flat precision $O(\pi^N),$ we can obtain an exact $\Delta$ and $P,Q$ 
known at precision at least $O(\pi^{N-cond(M)})$ with coefficients in $O_K$
and determinant in $O_K^*$ realizing an SNF form of $M.$
There is no loss in precision when computing $P^{-1}$ and $Q^{-1}.$
Now, computing $\Delta^{-1}$ exactly, its smallest coefficient being  of valuation $-cond(M),$
we obtain $M^{-1}=Q^{-1} \Delta^{-1} P^{-1}$ known at precision at least $O(\pi^{N-2cond(M)}),$
which concludes the proof.
\end{proof}

@@

From now on, we fix a matrix $M \in M_n(K)$ for which $d \chi_M$ is 
surjective. Let $(\alpha, P, Q, \chi_M)$ be the quadruaple encoding
the short form of $d \chi_M$; we recall that they are related by the
relations:
\begin{align*}
d \chi_M(dM) & =\tr(\com(X{-}M) \cdot dM) \\
\com(X{-}M) & = \alpha \cdot P V^t \cdot V Q^t \mod \chi_M.
\end{align*}
The matrix $P$ can be computed as follows. Pick $c \in K^n$. Define 
$c_i = M^i c$ for all $i \geq 1$. The $c_i$'s can be computed in 
$\softO(n^\omega)$ operations in $K$ using \cite{}. Let $P_\inv$ be the 
$n \times n$ matrix whose rows are the $c_i$'s for $1 \leq i \leq n$. 
Remark that $P_\inv$ is invertible if and only if $(c_0, c_1, \ldots, 
c_{n-1})$ is a basis of $K^n$ if and only if $c$ is a cyclic vector. 
Moreover after base change to the basis $(c_0, \ldots, c_{n-1})$, the matrix 
$M$ takes the shape \eqref{eq:companion}. In other words, if $P_\inv$
is invertible, then $P = P_\inv^{-1}$ is a solution of $M = P N P^{-1}$.
Observe moreover that the condition ``$P_\inv$ is invertible'' is open
for the Zariski topology. It then happens with high probability as soon
as it is not empty, that is as soon as $M$ admits a cyclic vector, which
holds by assumption.

>>>>

\newcommand{\noopsort}[1]{}
\DeclareMathOperator{\NP}{NP}
\DeclareMathOperator{\HP}{HP}
\DeclareMathOperator{\PP}{PP}
\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\val}{val}
\DeclareMathOperator{\pr}{pr}
\DeclareMathOperator{\tr}{Tr}
\DeclareMathOperator{\com}{Com}
\DeclareMathOperator{\Grass}{Grass}
\DeclareMathOperator{\Lat}{Lat}
\DeclareMathOperator{\round}{round}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\lcm}{lcm}
\DeclareMathOperator{\rec}{rec}
\DeclareMathOperator{\cond}{cond}

@@

\begin{prop}
Let $\cond(M)$ be the valuation of the last invariant factor of $M.$
If $dM$ is a flat precision of $O(\pi^N)$ on $M$ then $M^{-1}$
can be computed at precision $O(\pi^{N-2\cond(M)})$ by a \textbf{SNF} computation
and this lower-bound is optimal
at least when $N$ is large.
\end{prop}
\begin{proof}
The smallest valuation of a coefficient of $M^{-1}$ is $-\cond(M).$
It is $-2\cond(M)$ for $M^{-2}$ and it is then clear that $-2\cond(M)$
can be obtained as the valuation of a coefficient of $du(M) \cdot dM$
and the smallest that can be achieved this way for $dM$ in a precision lattice
of flat precision. Hence the optimality of the bound given, at least when 
$N$ is large, thanks to Lemma 3.14 of \cite{caruso-roe-vaccon:14a}

Now, the computation of an SNF form over a CDVF was described in \cite{Vaccon-these}.
From $M$ known at flat precision $O(\pi^N),$ we can obtain an exact $\Delta$ and $P,Q$ 
known at precision at least $O(\pi^{N-\cond(M)})$ with coefficients in $O_K$
and determinant in $O_K^*$ realizing an SNF form of $M.$
There is no loss in precision when computing $P^{-1}$ and $Q^{-1}.$
Now, computing $\Delta^{-1}$ exactly, its smallest coefficient being  of valuation $-\cond(M),$
we obtain $M^{-1}=Q^{-1} \Delta^{-1} P^{-1}$ known at precision at least $O(\pi^{N-2\cond(M)}),$
which concludes the proof.
\end{proof}

@@

From now on, we fix a matrix $M \in M_n(K)$ for which $d \chi_M$ is 
surjective. Let $(\alpha, P, Q, \chi_M)$ be the quadruaple encoding
the short form of $d \chi_M$; we recall that they are related by the
relations:
\begin{align*}
d \chi_M(dM) & =\tr(\com(X{-}M) \cdot dM) \\
\com(X{-}M) & = \alpha \cdot P V^t \cdot V Q^t \mod \chi_M.
\end{align*}
The matrix $P$ can be computed as follows. Pick $c \in K^n$. Define 
$c_i = M^i c$ for all $i \geq 1$. The $c_i$'s can be computed in 
$\softO(n^\omega)$ operations in $K$ using the first algorithm
of \cite{keller-gehrig:85a}. Let $P_\inv$ be the 
$n \times n$ matrix whose rows are the $c_i$'s for $1 \leq i \leq n$. 
Remark that $P_\inv$ is invertible if and only if $(c_0, c_1, \ldots, 
c_{n-1})$ is a basis of $K^n$ if and only if $c$ is a cyclic vector. 
Moreover after base change to the basis $(c_0, \ldots, c_{n-1})$, the matrix 
$M$ takes the shape \eqref{eq:companion}. In other words, if $P_\inv$
is invertible, then $P = P_\inv^{-1}$ is a solution of $M = P N P^{-1}$.
Observe moreover that the condition ``$P_\inv$ is invertible'' is open
for the Zariski topology. It then happens with high probability as soon
as it is not empty, that is as soon as $M$ admits a cyclic vector, which
holds by assumption.
